---
title: "Well-being comparison with wellber"
author: "John Vard"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The _wellber_ package is a wrapper for the [OECD's 2017 Better Life data](http://www.oecdbetterlifeindex.org/) and enables country well-being comparison across 24 Better Life indicators for 38 OECD countries.

## Basic functionality
The package provides simple _loading_, _selection_, and _plotting_ of the OECD's 2017 Better Life data, based on __country__, __indicator__, __sample type__, and a __window size__.

For ease-of-use the package includes a shiny app to allow more interactive exploration of a country's well-being performance. For those interested in exploring the app, call `wellber::launch.wellber()` to launch the _Well-Being Comparison Tool_. A screenshot of the tool is provided below.

```{r, echo=FALSE, out.width = "690px"}
knitr::include_graphics("/Users/jvard/Documents/MSc Data Analytics/Year_3/Adv Data Prog with R (STAT40830)/Wk_11/wellber/vignettes/wellber_app_screen.png")
```

## Using `load.wellber()`

`load.wellber()` loads the OECD dataset (stored internally within the package), removes unwanted information and creates the _Class_ column variable to group well-being indicators across 11  well-being categories.

```{r, warning=FALSE, message=FALSE}
library(wellber)
data_object = load.wellber()
head(data_object)
```

The [OECD's 2017 Better Life dataset](https://stats.oecd.org/index.aspx?DataSetCode=BLI) comprises of data from 38 countries, covering 24 well-being indicators. The raw dataset is composed of 8 variables and 3,388 individual data-points.

## Using `selector()`

`selector()` accepts a object of type _wellber_, as provided by `load.wellber()` and the function filters the data. 

For the below example a country is selected along with an well-being indicator. In this case, we're interested in exploring Chile's relative performance in relation to Housing Expenditure. The `samp_type` and `wind_size` parameters determine the sample type (_Men_, _Women_, _Total_) and plot window size (_Large_, _Medium_, _Small_).

```{r, warning=FALSE, message=FALSE}
data_object = load.wellber()
subset_data = selector(data_object, country = 'Chile', qual_ind = 'Housing expenditure', samp_type = 'Total', wind_size = 'Small')
subset_data
```

For guidance on the `country` and `qual_ind` options refer to the __Country and well-being list__ section at the bottom of this vingnette.

`wellber::selector()` also returns a list object of type _wellber_ that includes: 

1. A filtered tibble matching the arguments passed for a specific country, indicator and sample type is returned.
2. A tibble summary table relating to the indicator class performance for the choosen country and indicator.
3. A list of the min and max indicator values (for use during plotting).

## Using `plot()`

The filtered data from `selector()` may be plotted using `plot()`.

```{r, eval=FALSE}
data_object = load.wellber()
subset_data = selector(data_object, country = 'Chile', qual_ind = 'Housing expenditure', samp_type = 'Total', wind_size = 'Small')
plot(subset_data, ind_class = 0)
plot(subset_data, ind_class = 1)
```

```{r, echo=FALSE, out.width = "750px"}
knitr::include_graphics("/Users/jvard/Documents/MSc Data Analytics/Year_3/Adv Data Prog with R (STAT40830)/Wk_11/wellber/vignettes/wellber_plots.png")
```

The plots display the Housing expenditure for Chile (left) and housing performance generally (right) that is a combination of other housing related indicators. 

The _wind_size_ argument from `selector()` controls the plot window - in this case, only 5 bars are displayed. For each plot relative rankings are displayed above the bars to indicate performance across the 38 countries for that particular indicator. For the well-being indicator, minimum and maximum performance values are displayed as dashed lines on the plot.

### Country and well-being list

The `selector()` input arguments for `country` and `qual_ind` are covered below: 

Country | Well-Being Indicator
------------- | -------------
Australia | Dwellings without basic facilities 
Austria | Educational attainment
Belgium | Employees working very long hours
Brazil | Employment rate
Canada | Feeling safe walking alone at night
Chile | Homicide rate
Czech Republic | Household net adjusted disposable income
Denmark | Household net financial wealth
Estonia | Housing expenditure
Finland | Labour market insecurity
France | Life expectancy
Germany | Life satisfaction
Greece | Long-term unemployment rate
Hungary | Personal earnings
Iceland | Quality of support network
Ireland | Rooms per person
Israel | Self-reported health
Italy | Stakeholder engagement for developing regulations
Japan | Student skills
Korea | Time devoted to leisure and personal care
Latvia | Voter turnout
Luxembourg | Water quality
Mexico | Years in education
Netherlands |
New Zealand |
Norway |
Poland |
Portugal |
Russia |
Slovak Republic |
Slovenia |
South Africa |
Spain |
Sweden |
Switzerland |
Turkey |
United Kingdom |
United States |
